---
title: "Formative Assessment 9"
author: "Vera Aguila and Lindsy Masicat"
date: "2024-05-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1
A random variable _X_ is said to have the gamma distribution, or to be gamma distributed, if the density function is $f(x)= \left\{\begin{array}{lr} \frac{xÎ±^{-1}e^{-x/Î²}}{Î²^{Î±}Î“(Î±)} & x > 0 \\0 & x\leq 0 \end{array}\right.$  $(Î±,Î²>0)$, where  Î“(Î±) is the gamma function. Show that the mean and variance of the gamma distribution are given by (a) $ğœ‡=Î±Î²$, (b) $Ïƒ^{2}=Î±Î²^{2}$.

```{r}
alpha <- 2  
beta <- 3   

meanofgamma<- alpha * beta

varofgamma <- alpha * beta^2

cat("Mean (Î¼) of the gamma distribution:", meanofgamma, "\n")
cat("Variance (Ïƒ^2) of the gamma distribution:", varofgamma, "\n")

library(ggplot2)

x <- seq(0, 20, length.out = 1000)

pdf <- dgamma(x, shape = alpha, scale = beta)

mean_gamma <- alpha * beta
var_gamma <- alpha * beta^2

df <- data.frame(x = x, pdf = pdf)

ggplot(df, aes(x)) +
  geom_line(aes(y = pdf), color = "blue", size = 1) +
  geom_vline(xintercept = mean_gamma, linetype = "dashed", color = "red", size = 1) +
  annotate("text", x = mean_gamma + 1, y = max(pdf), label = paste("Mean (Î¼) =", round(mean_gamma, 2)), color = "red") +
  annotate("text", x = mean_gamma + 1, y = max(pdf) * 0.9, label = paste("Variance (Ïƒ^2) =", round(var_gamma, 2)), color = "blue") +
  labs(title = "Gamma Distribution", x = "x", y = "Probability Density") +
  theme_minimal()

```


## Problem 2
Prove that the mean and variance of a binomially distributed random variable are, respectively  $ğœ‡= np$ and and $Ïƒ^{2}=npq$.
 
```{r}
n <- 10    
p <- 0.5   

binomial_samples <- rbinom(10000, size = n, prob = p)

meanempirical <- mean(binomial_samples)
varempirical <- var(binomial_samples)

cat("Empirical Mean (Î¼) of the binomial distribution:", meanempirical, "\n")
cat("Empirical Variance (Ïƒ^2) of the binomial distribution:", varempirical, "\n")

meantheoretical <- n * p
vartheoretical <- n * p * (1 - p)

cat("Theoretical Mean (Î¼) of the binomial distribution:", meantheoretical, "\n")
cat("Theoretical Variance (Ïƒ^2) of the binomial distribution:", vartheoretical, "\n")

```



## Problem 3
Establish the validity of the Poisson approximation to the binomial distribution.

```{r}
library(ggplot2)

n <- 1000 
p <- 0.01 

# Generate data from binomial distribution
samplesofbinomial <- rbinom(10000, size = n, prob = p)

# Calculate the mean and variance of the binomial distribution
meanbinomial <- n * p
varbinomial <- n * p * (1 - p)

# Plot histogram of binomial samples
ggplot(data.frame(x = samplesofbinomial), aes(x)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black", alpha = 0.8) +
  ggtitle("Histogram of Binomial Distribution") +
  labs(x = "Value", y = "Frequency") +
  theme_minimal()

# Generate data from Poisson distribution with the same mean
lambda <- meanbinomial
poissonsamples <- rpois(10000, lambda)

# Plot histogram of Poisson samples
ggplot(data.frame(x = poissonsamples), aes(x)) +
  geom_histogram(binwidth = 1, fill = "salmon", color = "black", alpha = 0.8) +
  ggtitle("Histogram of Poisson Distribution") +
  labs(x = "Value", y = "Frequency") +
  theme_minimal()

```



